Fast Fourier transformation of electromagnetic data for computationally expensive kernels

### 2021-04-03 ManuscriptTime.tex
Draft RevII

\documentclass[extra, camera,%
    onecolumn,   % comment for two columns
    referee,     % uncomment for double lines
    % final,       % uncomment for changes
]{gji}
\usepackage{timet}

@@

\added{BEGIN TEMPORARY ADDED v v v v}

\rtodo 1 \emph{The independence from spatial model complexity is in my
  opinion debatable, because of a potential issue with the accuracy of phase
  information. It’s my experience that a high degree of imaginary-field
  component accuracy is needed for a proper f-t transform. --
  I also needed a high level of numerical accuracy in the f-domain, that is,
  very fine meshes and tight solver tolerances. An aspect which in my opinion
  is not addressed sufficiently in the current manuscript.}: Add somewhere that
  high accurate results are required.

\rtodo 2 \emph{In my first revision, I actually did not make a request
  for making the literature review "more extensive". Instead, I wrote that the
  authors "should spend more time researching the EM literature that is related
  to the actual topic." As it was and still is, half of the introduction
  (current manuscript, page 2 to p. 3, line 46) is a rather broad/general EM
  overview. Given the fact that the paper \textbf{connects directly to Plessix
  et al. (2007) and Mulder et al. (2008)}, I would have expected a \textbf{more
  targeted introduction towards the specific advances of this manuscript} –
  instead of talking about a choice of only loosely connected and more general
  topics like Lebedev grids, tetrahedral grids, integral equation, direct
  solvers, etc. I like that the introduction is generally concise. My
  impression remains that the introduction can have the same length but be more
  targeted. Other readers may also see it that way. I won’t need another
  explanation about your potentially differing opinion. This is just a
  suggestion for improvement. Feel free to ignore.}

\rtodo 3 Reply; Include comments from earlier review; papers that were
referred; difficulties in late/early times; IP effects. Refer to steel-casing
effects (lindsey), but state we look here at non-Debye cole models.
Add also the papers seogi mentioned re IP effects. \cite{CAG.19.Heagy}
\cite{EXP.20.Kang} \emph{We tried to improve the introduction and focus more on
  the actual topic. Amongst other, we address the topic of steel casing and
  non-Debye Cole-Cole models.}

\added{END TEMPORARY ADDED n n n}

The controlled-source electromagnetic (CSEM) method is one of the common
non-seismic tools in exploration geophysics, not only in hydrocarbon
exploration \citep{GEO.10.Constable}, but also in the search for sulfides
\citep{GRL.19.Gehrmann}, water \citep{GEO.05.Pedersen}, geothermal sources
\citep{WGC.15.Girard}, or for geological purposes \citep{NAT.19.Johanson}.
While current sources with a few frequencies are used in the deep marine
environment, transient measurements are more common in the shallow marine
environment and on land \citep[e.g., ][]{GEO.07.Ziolkowski, SEG.07.Andreis,
SEG.07.Avdeeva}. One of the main reasons is the dominance of the airwave in
shallow marine and terrestrial measurements, which can be better separated in
the time domain. CSEM is usually divided into frequency- and time-domain
methods, depending on whether the source signal is a continuous waveform, such
as a sine, or a finite waveform, such as a pseudo-random binary sequence
(PRBS). A numerical comparison of the two methods is given by
\cite{GP.13.Conell}. Acquired CSEM data are subsequently often analysed
(modelled and inverted for) in their respective domain, either the frequency or
time domain. Modellers of layered media usually exploit the horizontal
shift-invariance by computing the responses in the wavenumber-frequency domain
followed by a 2D inverse spatial Fourier transform, also called Hankel
transform, to the space-frequency domain, and a regular inverse Fourier
transform if time-domain data are required \citep[e.g., ][]{GEO.15.Hunziker}.

CSEM codes for 2D and 3D computations, on the other hand, mostly compute their
responses directly in the required domain, either frequency or time. There is a
wealth of 3D electromagnetic codes, with a first boom in the 1990's;
\cite{B.SEG.99.Oristaglio} provide a good review of the state of art at that
time. A second wave occurred during the CSEM boom in the hydrocarbon industry
in the early 21th century, see, e.g., \cite{SG.05.Avdeev} and
\cite{SG.10.Borner} for comprehensive overviews. Recent publications after
these review articles span the width of numerical methods from integral
equation codes \citep{MGS.17.Kruglyakov} and differential equation methods such
as finite differences \citep{GEO.10.Mittet, GEO.11.Zaslavsky, CAG.13.Sommer},
finite elements \citep{GJI.13.Grayver}, and finite volumes
\citep{GEO.14.Jahandari}. Besides the different methods there are also many
different types of discretisation. Commonly used, besides rectilinear Yee grids
\citep{IEEE.66.Yee} as used by our code, are Lebedev grids
\citep{CMMP.64.Lebedev} but also unstructured grids using, for instance,
tetrahedra \citep{CAG.17.Cai}. The computational requirement to model CSEM data
also depends heavily on the technique used to solve the system. There are
direct solvers \citep{GEO.15.Grayver}, indirect solvers
\citep{GJI.15.Jaysaval}, and more recently also hybrid solvers
\citep{GEO.18.Liu}.

Under certain conditions it can be a viable alternative to model time-domain
data with a frequency-domain code, as shown by \cite{GEO.08.Mulder}. They
compared the computational complexity of explicit time-stepping schemes (Du
Fort-Frankel method as used in, e.g., \cite{GEO.04.Commer, GEO.07.Maao});
implicit schemes \citep[e.g., ][]{GEO.04.Haber}, and matrix exponentials and
Lanczos reduction schemes \cite[e.g., ][]{RS.94.Druskin} with the Fourier
transform of frequency-domain solutions. Their conclusion was that the Fourier
transform method can be favourable if the number of required frequencies is
small relative to the number of grid points in each coordinate. The conditions
for fast computation of time-domain data with a frequency-domain code are: a
sufficiently powerful solver, appropriate frequency selection and
interpolation, and an automated gridding, for which they used the
multi-frequency CSEM approach presented by \cite{GEO.07.Plessix}. We build upon
these results but improve the run time from hours to minutes. The main reasons
for this significant speed-up are an adaptive, frequency-dependent gridding
scheme that minimizes the required cells in each dimension, and a logarithmic
Fourier transform such as digital linear filters \citep[DLF, ][]{GP.71.Ghosh}
or the logarithmic fast Fourier transform \citep[FFTLog, ][]{RAS.00.Hamilton}
to go from the frequency to the time domain. The latter makes it also possible
to only use the imaginary part of the frequency-domain response, which has
advantages when it comes to interpolation.

@@

\rtodo \added{REDO and send to R.R}

\rtodo \added{Add citation to R.R. (2021) regarding RA and speed-up; limitation
to switch-on/-off.}

\rtodo \added{Add notion about preference for f-domain for IP problems}

\rtodo \added{Extra example: No-IP vs IP; land, 1D}

@@

We present improvements to both the adaptive gridding and the transform from
frequency domain to time domain, which makes time-domain modelling with a
frequency-domain code viable in comparison with time-domain codes. It is
important to note that while we show the transform from frequency to time
domain, the conclusions regarding the Fourier transform can equally well be
applied to the transform from time-domain to frequency-domain data for a code
that solves Maxwell's equations in the time domain. \deleted{A variation of our
method was implemented, for instance, by \custem \mbox{\citep{GEO.19.Rochlitz}}
to obtain frequency-domain data from time-domain computations (Rochlitz,
private communication, 2020).}

@@

\added{The proposed Fourier transformation requires a robust solver that can
compute precise results over a wide range of frequencies (times) to obtain
time-domain (frequency-domain) responses. In our case the solver uses a
regular, stretched grid and computes the electric fields in the frequency
domain. This setup is the target of the following gridding recommendations.
Whilst they will look differently for other mesh types or a time-domain code,
some conclusion will still hold.}

@@

\todo[inline]{HERE HERE HERE}

@@

Figure~\ref{fig:marine} shows the result for an offset of 5\,km, in (a) the
frequency and (b) the time domain. The recovered response with the 3D code
captures the airwave (first peak) and the subsurface (second peak) very
accurately. At later times the error starts to increase. We would need to
compute a few additional lower frequencies if we want to improve it.
%
\begin{figure*}
  \centering
  \includegraphics[width=\fwidth]{06-marine}
  \caption{Electric inline response at an offset of 5\,km for a shallow marine,
    layered scenario. (a) Frequency-domain response, where the blue circles
    denote computed responses and the black dots interpolated responses or
    responses set to zero. (b) Time-domain response.}
  \label{fig:marine}
\end{figure*}
%
In the frequency-domain plot, it can be seen that the high frequencies are not
computed very accurately, but without too much influence on the time-domain
response. These frequencies could be left out if an offset of 5\,km is the only
objective. However, we also want to retrieve shorter offsets from the same
computation, for which these frequencies are required.

@@

\added{$\Downarrow\qquad$ Note to reviewers: The entire section 4.3 below is
new, without being all marked in green. $\qquad\Downarrow$}

\added{Mention independence of spatial complexity, why we also show 1D
examples; however, sufficiently accurate 3D results are an obvious
pre-requisite.}

The Cole-Cole model \citep[CCM, ][]{JCP.41.Cole}, written in terms of
conductivities instead of electric permittivities as by the original authors,
is given by
%
\begin{equation}
  \sigma(\omega) = \sigma_\infty + \frac{\sigma_0 - \sigma_\infty}{1 +
    (\rm{i}\omega\tau)^c}\ ,
  \label{eq:ccm}
\end{equation}
%
where $\sigma_0$ and $\sigma_\infty$ refer to the low-frequency and
high-frequency conductivity values, respectively, $\tau$ is the central
relaxation time, and $c$ the CCM exponent describing the broadness of the
relaxation time distribution. Note that this model slightly differs from the
one phrased in terms of resistivities given by \cite{GEO.78.Pelton}, see, e.g.,
\cite{GJI.13.Tarasov}.
%
\begin{figure}
  \centering
  \includegraphics[width=\fwidth]{11-cole-cole-model}
  \caption{}
  \label{fig:CCM}
\end{figure}
%

It required only 30 frequencies from 1.5e-8 to 8.3\,Hz, instead of the 244
frequencies from 1.5e-8 to 6.8e6\,Hz required by the digital filter (201\,pt
filter from Key, 2012). Besides the fact that the responses for the higher
frequencies can be set to zero it is also sufficient to only compute every
5th of the required frequencies.

low frequencies cannot be 'interpolated'; however, other tricks could be
implemented. The DC values (so low frequencies or late times) is the same
independent of the IP effect. This makes it possible to estimate the DC value
with the IP-free model which does not require very low frequencies, and use
consequently this DC value to interpolate the low frequencies.

for c>0.6, 17 frequencies from 1.2e-4 to 8.3\,Hz would be sufficient.

%    'src': (0, 0, 0.01),
%    'rec': (900, 0, 0.01),
%    'depth': [0, 300, 400],
%    res = 2e14, 1, 1, 1
%    layer: cond_8 = 1.25, cond_0 = 1 => m = 0.2
%    tau: 1
%    fmin=1e-4,
%    fmax=5e2,
%    signal=1,        # <= Step-On!
%    every_x_freq=6,  # Only every 6th value

%  time        [s] :  0.01 - 10000 : 301  [min-max; #]
%  Fourier         :  DLF (Sine-Filter)
%    > Filter      :  Key 601 CosSin (2009)
%    > DLF type    :  Lagged Convolution
%  Req. freq  [Hz] :  6.67463E-18 - 4.00754E+13 : 747  [min-max; #]
%  Calc. freq [Hz] :  0.000155601 - 424.87 : 27  [min-max; #]

% Comparison: Full DLF; cosine-filter step-off, 747 frequencies
% Our method: 27 freqs; DC(no IP) - step-on; DC at 1000 s
%    (no more frequencies were required)
%    important to see that the DC value (low freq, late t; is the same IP/noIP)

The actual transform was carried out with the 601-point sine-cosine filter from
\cite{GEO.09.Key}.

Key 601\,pt filter is very long; seems to be by far the best for low c values.
Designing a new filter specific for IP problems might be the best approach

DC: 1000s; the same amount of frequencies is required to compute times up to
10s or up to 1000s

\cite{GJI.17.Commer}

\begin{itemize}
  \item Designing a DLF for IP effects
  \item DC value is the same for the IP or non-IP model, so one could use
    the non-IP response and do interpolation
\end{itemize}

\added{$\Uparrow\qquad$ Note to reviewers: The entire section 4.3 above is new,
without being all marked in green. $\qquad\Uparrow$}

@@

Figure~\ref{fig:marine-wrong-x-y} shows this effect. It is the same model as in
the previous section; however, for the adaptive gridding in the horizontal
directions, $\rho_\mr{ave}=1\,\ohmm$ was used instead of
$\rho_\mr{ave}=\num{10000}\,\ohmm$. Having the boundaries too near in the
horizontal directions leads to worse results for most frequencies and entirely
wrong results for high frequencies. Comparison with the 1D result in the time
domain shows that it is the airwave whose amplitude is heavily overestimated.
%
\begin{figure*}
  \centering
  \includegraphics[width=\fwidth]{08-marine-wrong-x-y}
  \caption{Same model as used for Figure~\ref{fig:marine}, but with the
    horizontal boundaries not far enough. Although the resulting time-domain
    result looks plausible, the comparison with the 1D result shows that it
    significantly overestimates the amplitude of the airwave.}
  \label{fig:marine-wrong-x-y}
\end{figure*}
%
It can be difficult to spot these errors in the time-domain result, as the
response looks plausible. Only a comparison with the 1D result reveals that it
is actually wrong. A possibility to detect such problems for complicated cases,
where there is no semi-analytical result to compare with, is to compute two or
more models, moving the boundary. When the responses stop to change, one can
assume that the boundary is far enough. Another possibility is to look at the
amplitudes close to the boundaries and ensure that they are small enough.

>>>>

\documentclass[extra, camera,%
    onecolumn,   % comment for two columns
    referee,     % uncomment for double lines
    final,       % uncomment for changes
]{gji}
\usepackage{timet}

@@

The controlled-source electromagnetic (CSEM) method is one of the common
non-seismic tools in exploration geophysics, not only in hydrocarbon
exploration \citep{GEO.10.Constable}, but also in the search for sulfides
\citep{GRL.19.Gehrmann}, water \citep{GEO.05.Pedersen}, geothermal sources
\citep{WGC.15.Girard}, or for geological purposes \citep{NAT.19.Johanson}.
While current sources with a few frequencies are used in the deep marine
environment, transient measurements are more common in the shallow marine
environment and on land \citep[e.g., ][]{GEO.07.Ziolkowski, SEG.07.Andreis,
SEG.07.Avdeeva}. One of the main reasons is the dominance of the airwave in
shallow marine and terrestrial measurements, which can be better separated in
the time domain. CSEM is usually divided into frequency- and time-domain
methods, depending on whether the source signal is a continuous waveform, such
as a sine, or a finite waveform, such as a pseudo-random binary sequence
(PRBS). A numerical comparison of the two methods is given by
\cite{GP.13.Conell}. Acquired CSEM data are subsequently often analysed
(modelled and inverted for) in their respective domain, either the frequency or
time domain. Modellers of layered media usually exploit the horizontal
shift-invariance by computing the responses in the wavenumber-frequency domain
followed by a 2D inverse spatial Fourier transform, also called Hankel
transform, to the space-frequency domain, and a regular inverse Fourier
transform if time-domain data are required \citep[e.g., ][]{GEO.15.Hunziker}.
CSEM codes for 2D and 3D computations, on the other hand, mostly compute their
responses directly in the required domain, either frequency or time.

\added{Electromagnetic methods in geophysics span a wide range of model
scenarios and acquisition layouts which have all different implications. A
well-known case is for instance the land or shallow marine case versus the deep
marine case, where in the former case a dominating airwave has to be accurately
modelled which can can be completely ignored in the latter case. Recent areas
of particular interest in time-domain modelling that pose numerical challenges
are, for instance, simulating the fields through steel-cased wells
\mbox{\citep{CAG.19.Heagy}} or the effects of induced polarization
\mbox{\citep{EXP.20.Kang}}. The former is challenging because of the high
conductivity contrasts requiring very detailed meshing, the latter is
challenging for time-domain modelling because the models are frequency
dependent.}

\deleted{There is a wealth of 3D electromagnetic codes, with a first boom in the
1990's; \mbox{\cite{B.SEG.99.Oristaglio}} provide a good review of the state of
art at that time. A second wave occurred during the CSEM boom in the
hydrocarbon industry in the early 21th century, see, e.g.,
\mbox{\cite{SG.05.Avdeev}} and \mbox{\cite{SG.10.Borner}} for comprehensive
overviews. Recent publications after these review articles span the width of
numerical methods from integral equation codes \mbox{\citep{MGS.17.Kruglyakov}}
and differential equation methods such as finite differences
\mbox{\citep{GEO.10.Mittet, GEO.11.Zaslavsky, CAG.13.Sommer}}, finite elements
\mbox{\citep{GJI.13.Grayver}}, and finite volumes
\mbox{\citep{GEO.14.Jahandari}}. Besides the different methods there are also
many different types of discretisation. Commonly used, besides rectilinear Yee
grids \mbox{\citep{IEEE.66.Yee}} as used by our code, are Lebedev grids
\mbox{\citep{CMMP.64.Lebedev}} but also unstructured grids using, for instance,
tetrahedra \mbox{\citep{CAG.17.Cai}}. The computational requirement to model
CSEM data also depends heavily on the technique used to solve the system. There
are direct solvers \mbox{\citep{GEO.15.Grayver}}, indirect solvers
\mbox{\citep{GJI.15.Jaysaval}}, and more recently also hybrid solvers
\mbox{\citep{GEO.18.Liu}}.}

Under certain conditions it can be a viable alternative to model time-domain
data with a frequency-domain code, as shown by \cite{GEO.08.Mulder}. They
compared the computational complexity \added{of different time-domain methods
with the computational complexity of frequency-domain computation followed by a
Fourier transform:} \deleted{of} explicit time-stepping schemes (Du
Fort-Frankel method as used in, e.g., \cite{GEO.04.Commer,
GEO.07.Maao})\replaced{ and}{;} implicit schemes \citep[e.g.,
][]{GEO.04.Haber}\added{ have a complexity of $O(n^4)$}, \deleted{and} matrix
exponentials and Lanczos reduction schemes \cite[e.g., ][]{RS.94.Druskin}
\added{have a complexity of $O(n^4\sqrt{\log n})$}\replaced{, and}{with the}
Fourier transform of frequency-domain solutions \added{have a complexity $O(n_f
n^3)$}; \added{$n$ is the number of grid points in each direction, and $n_f$
the number of frequencies}. Their conclusion was that the Fourier transform
method can be favourable if the number of required frequencies is small
relative to the number of grid points in each coordinate. The conditions for
fast computation of time-domain data with a frequency-domain code are: a
sufficiently powerful solver, appropriate frequency selection and
interpolation, and an automated gridding, for which they used the
multi-frequency CSEM approach presented by \cite{GEO.07.Plessix}. We build upon
these results but improve the run time from hours to minutes. The main reasons
for this significant speed-up are \added{the further reduction of required
frequencies by introducing lower and upper thresholds of numerically important
frequencies,} an adaptive, frequency-dependent gridding scheme that minimizes
the required cells in each dimension, and a logarithmic Fourier transform such
as digital linear filters \citep[DLF, ][]{GP.71.Ghosh} or the logarithmic fast
Fourier transform \citep[FFTLog, ][]{RAS.00.Hamilton} to go from the frequency
to the time domain. The latter makes it also possible to only use the imaginary
part of the frequency-domain response, which has advantages when it comes to
interpolation.

@@

@@

We present improvements to both the adaptive gridding and the transform from
frequency domain to time domain, which makes time-domain modelling with a
frequency-domain code viable in comparison with time-domain codes. It is
important to note that while we show the transform from frequency to time
domain, the conclusions regarding the Fourier transform can equally well be
applied to the transform from time-domain to frequency-domain data for a code
that solves Maxwell's equations in the time domain. \added{Our recommendations
for speeding up the Fourier transform for expensive kernels are independent of
spatial complexity, as diffusive EM responses are smooth functions of both
frequency and time. However, the intrinsic requirement is an accurate solver
over a sufficient wide range of frequencies or times.} \deleted{A variation of
our method was implemented, for instance, by \custem
\mbox{\citep{GEO.19.Rochlitz}} to obtain frequency-domain data from time-domain
computations (Rochlitz, private communication, 2020).}

@@

\added{The proposed Fourier transformation requires a robust solver that can
compute accurate results over a wide range of frequencies (times) to obtain
time-domain (frequency-domain) responses. In our case the solver uses a
regular, stretched grid and computes the electric fields in the frequency
domain. This setup is the target of the following gridding recommendations.
Whilst they will look differently for other mesh types or a time-domain code,
some conclusion will still hold.}

@@

@@

Figure~\ref{fig:marine} shows the result for an offset of 5\,km, \added{(a)} in
\deleted{(a)} the frequency \added{domain} and (b) \added{in} the time domain.
The recovered response with the 3D code captures the airwave (first peak) and
the subsurface (second peak) very accurately. At later times the error starts
to increase. We would need to compute a few additional lower frequencies if we
want to improve it.
%
\begin{figure*}
  \centering
  \includegraphics[width=\fwidth]{06-marine}
  \caption{Electric inline response at an offset of 5\,km for a shallow marine,
    layered scenario. (a) Frequency-domain response, where the blue circles
    denote computed responses and the black dots interpolated responses or
    responses set to zero. (b) Time-domain response.}
  \label{fig:marine}
\end{figure*}
%
In the frequency-domain plot, it can be seen that the high frequencies are not
computed very accurately, but without too much influence on the time-domain
response. These frequencies could be left out if an offset of 5\,km is the only
objective. However, we also want to retrieve shorter offsets from the same
computation, for which these frequencies are required.

@@

@@

Figure~\ref{fig:marine-wrong-x-y} shows this effect. It is the same model as in
the previous section; however, for the adaptive gridding in the horizontal
directions, $\rho_\mr{ave}=1\,\ohmm$ was used instead of
$\rho_\mr{ave}=\num{10000}\,\ohmm$. Having the boundaries too near in the
horizontal directions leads to worse results for most frequencies and entirely
wrong results for high frequencies. Comparison with the 1D result in the time
domain shows that it is the airwave whose amplitude is heavily overestimated.
%
\begin{figure*}
  \centering
  \includegraphics[width=\fwidth]{08-marine-wrong-x-y}
  \caption{Same model as used for Figure~\ref{fig:marine}, but with the
    horizontal boundaries not far enough. Although the resulting time-domain
    result looks plausible, the comparison with the 1D result shows that it
    significantly overestimates the amplitude of the airwave.}
  \label{fig:marine-wrong-x-y}
\end{figure*}
%
It can be difficult to spot these errors in the time-domain result, as the
response looks plausible. Only a comparison with the 1D result reveals that it
is actually wrong. A possibility to detect such problems for complicated cases,
where there is no semi-analytical result to compare with, is to \replaced{carry
out a mesh-convergence test by computing}{compute} two or more models, moving
the boundary. When the responses stop to change, one can assume that the
boundary is far enough. Another possibility is to look at the amplitudes close
to the boundaries and ensure that they are small enough.

\added{The proposed Fourier-transform approach works independently of spatial
complexity, as diffusive electromagnetic fields are smooth functions of both
frequency and time. An interesting test it to see whether this applies as
well for dispersive media where the model parameters are frequency dependent,
such as is the case in induced polarization. To test this we use the Cole-Cole
model \mbox{\citep[CCM, ][]{JCP.41.Cole}} which, written in terms of
conductivities instead of electric permittivities as by the original authors,
is given by}
%
\begin{equation}
\added{
  \sigma(\omega) = \sigma_\infty + \frac{\sigma_0 - \sigma_\infty}{1 +
    (\rm{i}\omega\tau)^c}\ ,
}
  \label{eq:ccm}
\end{equation}
%
\added{where $\sigma_0$ and $\sigma_\infty$ refer to the low-frequency and
high-frequency conductivity values, respectively, $\tau$ is the central
relaxation time (s), and $c$ the CCM exponent describing the broadness of the
relaxation time distribution. Note that this model slightly differs from the
one phrased in terms of resistivities given by \mbox{\cite{GEO.78.Pelton}},
see, e.g., \mbox{\cite{GJI.13.Tarasov}}. }

\added{\mbox{\cite{GJI.17.Commer}}, have shown that time-domain IP modelling
can be difficult, particularly for low $c$-values. Modelling
frequency-dependent models in the frequency domain is for any given frequency
naturally no different than modelling a non frequency-dependent model. However,
the interesting point is to see how the Fourier transform behaves to obtain
time-domain data from frequency-domain responses. As an example we modelled the
response for a land case, air above a halfspace of 1\,S/m. Within the halfspace
is a dispersive, 100\,m thick layer at a depth of 300\,m, with
$\sigma_0=1.0\,$S/m and $\sigma_\infty=1.25\,$S/m; $\tau=1$, and we vary
$c$ from 0 (no IP) to 1.0, 0.75, 0.5, and 0.25. The x-directed step-off source
and the receiver at an inline offset of 900\,m are place at the surface. The
real and imaginary frequency-domain responses for this model are shown in the
left column of Figure~\ref{fig:CCM}. It can be seen that the effect of the
IP layer in the frequency domain is that the real part starts to deviate from
the DC value at lower frequencies. The imaginary part decays slower towards
zero when decreasing the frequency, and it deviates from a linear function of 
frequency (on log-log scale) at lower frequencies as well. This indicates that
we need lower frequency content for the Fourier transform the smaller $c$
becomes.}
%
\begin{figure}
  \centering
  \includegraphics[width=\fwidth]{11-cole-cole-model}
  \caption{\added{IP example for different $c$ values, frequency-domain
  response on the left and time-domain response on the right. The frequencies
  indicated by the blue vertical bars in the lower-left figure were used for
  our Fourier transform approach, requiring only the computation of 27
  frequencies instead of the full 747 frequencies required by the digital
linear filter method.}}
  \label{fig:CCM}
\end{figure}
%

\added{For the Fourier transform we used the 601-point sine-cosine filter from
\mbox{\cite{GEO.09.Key}}. This is a very long filter, but it seemed to be by
far the best filter for low $c$-values. However, designing a new filter
particularly for IP problems might be a better approach. The selected frequency
thresholds were chosen at $f_\mr{min}=10^{-4}\,$Hz and $f_\mr{max}=500\,$Hz,
and within the thresholds we only computed every 6th frequency. Using these
values only 27 frequencies have to be computed, instead of 747 frequencies
required by the actual transform. To obtain the step-off response we computed
the step-on response and subtracted it from the DC value at 1000\,s (without
the need for any additional frequencies).}
